\Large \bf Apropos Replacement : Development of a full text search tool for man pages

### 2012-02-29 apropos_replacement.tex
Put the listings together on a separate page using single column layout.

\subsection{Lack of support for free form queries}
As noted earlier the conventional \textit{apropos} is limited to the keywords
used in the NAME section of the man pages. This means that the users do not have
a whole lot of options for specifying keywords in their queries. Besides this,
they cannot search for keywords which are not usually specified in the title of
a man page, for example a query like {\tt EINVAL} would return no results. \\

{\small
\lstset{language = XML, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%add new user%\normalfont\upshape''%
add new user: nothing appropriate

@@

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it is not smart enough to provide support for basic natural language processing
constructs,
like stemming or spelling corrections.
For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query
\textit{``make directories''}, it fails when the keyword \textit{``directory''}
is used in place of \textit{``directories''}, even though the two words are based
on same root word. \\

{\small
\lstset{language = XML, caption={No support for stems or words with same root},}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%make directories%\normalfont\upshape"%
mkdir (1) - make directories

@@

Similarly, it is very common for users to misspell keywords in a query and \textit{apropos} has no support for it either.

{\tt \small
\lstset{language = sh, caption={No spelling correction},}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%copy strings%\normalfont\upshape''%
stpcpy, stpncpy, strcpy, strncpy (3) - copy
strings

@@

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes its output can be
unintelligible and it can be very hard for the user to identify relevant results. \\

{\tt \small
\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ apropos power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
}

@@

\begin{description}
\item[Database Schema] \hfill \\
During the initial stage of the project, the structure of the database was kept
simple. The FTS table consisted of 3 columns:
{\tt name}, {\tt name\_desc}, and {\tt desc}.
The
{\tt
name
}
and
{\tt
name\_desc
}
columns stored the name of the man page and the one line description from the
{\tt
NAME
}
section respectively, while the
{\tt
desc
}
column stored the content from the rest of the sections. This approach worked
quite well for a small set of man pages (for example the 4000 or so man pages in
the base set of NetBSD). However, as the number of documents were increased (for example including man pages from pkgsrc), the quality of search started deteriorating. This was mainly because bulk of the parsed data was being stored
in the same column in the database and the ranking algorithm had no way of
identifying more relevant results except from identifying whether the match was found in the {\tt NAME} section or any other section. \\

To rectify this, first there was an attempt to try out better ranking schemes
(the various ranking schemes that were tried out during the development of the
project are discussed later). Better ranking schemes required storing precomputed
term weights [19] on the file system, this almost doubled the storage requirements, therefore such ranking schemes were discarded later on, and the
database schema was decomposed further into more columns. The decomposition into
more columns was done to represent the most commonly occurring sections in usual Unix man pages, for example
{\tt NAME}, {\tt DESCRIPTION}, {\tt LIBRARY}, {\tt RETURN VALUES},
{\tt EXIT STATUS}, {\tt ERRORS}
etc. Such a decomposition allowed for an elaborate ranking scheme, which could
assign different weights to each of these columns, based on their relevancy. The
final schema of the database is shown in {\bf Listing 5}
\end{description}

@@

{\tt \small
\lstset{language = XML, caption={Database Schema},}
\begin{lstlisting}
The Database has three tables at present:
1. mandb
2. mandb_meta
3. mandb_links

  COLUMN NAME       DESCRIPTION
  1. section        Section number
  2. name           The name of the page  
  3. name_desc      Short description from
                    NAME section  
  4. desc           Contains the content
                    from DESCRIPTION
                    section and any other
                    section which is not
                    stored in a separate
                    column.
  5. lib            The LIBRARY section
  6. return_vals    RETURN VALUES section
  7. env            ENVIRONMENT section
  8. files          FILES section
  9. exit_status    EXIT STATUS section
  10. diagnostics   DIAGNOSTICS section
  11. errors        ERRORS section
  12. md5_hash      An MD5 hash of the
                    contents of the man
                    page (UNIQUE).
(2) mandb_meta:
    This table maintains essential
    metadata of all the indexed
    man page files.

  COLUMN NAME       DESCRIPTION
  1. device         (dev_t)Logical device
                    number from stat(2)
  2. inode          (ino_t)Inode number
                    from stat(2)
  3. mtime          Last modification time
                    from stat(2)
  4. file           Absolute path name
                    (UNIQUE constraint)
  5. md5_hash       MD5 Hash of the man
                    page
  6. id             A unique integer ID
                    referring to a page
                    in mandb (PRIMARY KEY)

(3) mandb_links:
    This is an index of all the hard/soft
    links of the man pages. The list of
    the linked pages is usually obtained 
    from the multiple .Nm entries in the
    page.

  COLUMN NAME       DESCRIPTION
  1. link           The name of the hard/
                    soft link (UNIQUE index) 
  2. target         Name of the target
                    page to which it
                    points  
  3. section        The section number  
  4. machine        The machine
                    architecture (if any)
                    for which the page is
                    relevant.
\end{lstlisting}
}

@@

\section{Results}
This section shows results of some of the sample queries on new \textit{apropos}
to demonstrate how it performs as compared to the traditional version. \\
%\footnote{Although the new \textit{apropos} by default returns 10
%results for a query but in the listings the output has been
%snipped to save space
%} \\

{ \small
\lstset{language = XML, caption={Add new user},}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%add new user%\normalfont\upshape''%
ssh-add(1)      adds private key identities to
the authentication agent...on the command line.
If any file requires a passphrase, ssh-add asks
for the passphrase from the user. The
passphrase is read from the user's tty. ssh-add
retries the last passphrase if multiple
identity files are given...

@@

{\tt \small
\lstset{language = XML, caption={make directory},}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%make directory%\normalfont\upshape''%
make(1) maintain program dependencies
...CURDIR A path to the directory where make
was executed. Refer to the description of PWD
for more details. MAKE The name that make was
executed with argv[0].
For compatibility make also sets .MAKE with the
same value. The...

@@

{\tt \small
\lstset{language = XML, caption={signal number to string},}
\begin{lstlisting}
$ apropos %\normalfont\upshape``%signal number to string%\normalfont\upshape''%
psignal(3)      system signal messages
...the signal number is not recognized
sigaction(2) , the string Unknown signal
is produced. The psiginfo function produces
the same output as the psignal function,
only it uses the signal number information
from the si argument. The message strings can...

intro(2)        introduction to system calls
and error numbers
...undefined signal to a signal(3) or kill(2)
function). 23 ENFILE Too many open files in
system . Maximum number...shell. Pathname A
path name is a NUL -terminated character
string starting with an optional slash,
followed by zero or...

@@

\section{Acknowledgement}
This project has been developed as part of \emph{Google Summer of Code 2011}
[17], thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. Thanks to \textbf{David Young} who was involved with
this project closely and offered useful guidance and encouragement throughout. A
special thanks goes to \textbf{Thomas Klausner} who helped in writing and
reviewing the man pages for this project. Thanks to \textbf{Petra Zeidler} for
administering the GSoC program for \textit{The NetBSD Foundation}. \\

@@

\end{thebibliography}

>>>>

\subsection{Lack of support for free form queries}
As noted earlier the conventional \textit{apropos} is limited to the keywords
used in the NAME section of the man pages. This means that the users do not have
a whole lot of options for specifying keywords in their queries. Besides this,
they cannot search for keywords which are not usually specified in the title of
a man page, for example a query like {\tt EINVAL} would return no results.
See {\bf listing \ref{lst:noFreeFormQuery}} for concrete examples.  \\

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it is not smart enough to provide support for basic natural language processing
constructs,
like stemming or spelling corrections.
As it can be seen in {\bf listing \ref{noStemming}},
while \textit{apropos} returns the correct result for the query
\textit{``make directories''}, it fails when the keyword \textit{``directory''}
is used in place of \textit{``directories''}, even though the two words are based
on same root word. \\

Similarly, it is very common for users to misspell keywords in a query and \textit{apropos} has no support for detecting it and offering any useful behavior either (see {\bf listing \ref{noSpell}}).

\subsection{Unintelligible Output}
Because of the way \textit{apropos} works, sometimes its output can be
unintelligible and it can be very hard for the user to identify relevant results, see {\bf listing \ref{unintelligibleOutput}} for an example. \\

{\small
\lstset{language = XML, caption={No results for free form queries},
label={lst:noFreeFormQuery}}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%add new user%\normalfont\upshape''%
add new user: nothing appropriate

@@

{\small
\lstset{language = XML, caption={No support for stems or words with same root},
label=noStemming}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%make directories%\normalfont\upshape"%
mkdir (1) - make directories

@@

{\tt \small
\lstset{language = sh, caption={No spelling correction}, label=noSpell}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%copy strings%\normalfont\upshape''%
stpcpy, stpncpy, strcpy, strncpy (3) - copy strings

@@

{\tt \small
\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},
label=unintelligibleOutput}
\begin{lstlisting}[float=*]
$ apropos power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
}

@@

\begin{description}
\item[Database Schema] \hfill \\
During the initial stage of the project, the structure of the database was kept
simple. The FTS table consisted of 3 columns:
{\tt name}, {\tt name\_desc}, and {\tt desc}. The
{\tt
name
}and
{\tt
name\_desc
}columns stored the name of the man page and the one line description from the
{\tt
NAME
}section respectively, while the
{\tt
desc
}column stored the content from the rest of the sections. This approach worked
quite well for a small set of man pages (for example the 4000 or so man pages in
the base set of NetBSD). However, as the number of documents were increased (for example including man pages from pkgsrc), the quality of search started deteriorating. This was mainly because bulk of the parsed data was being stored
in the same column in the database and the ranking algorithm had no way of
identifying more relevant results except from identifying whether the match was found in the {\tt NAME} section or any other section. \\

To rectify this, first there was an attempt to try out better ranking schemes
(the various ranking schemes that were tried out during the development of the
project are discussed later). Better ranking schemes required storing precomputed
term weights [19] on the file system, this almost doubled the storage requirements, therefore such ranking schemes were discarded later on, and the
database schema was decomposed further into more columns. The decomposition into
more columns was done to represent the most commonly occurring sections in usual Unix man pages, for example
{\tt NAME}, {\tt DESCRIPTION}, {\tt LIBRARY}, {\tt RETURN VALUES},
{\tt EXIT STATUS}, {\tt ERRORS}
etc. Such a decomposition allowed for an elaborate ranking scheme, which could
assign different weights to each of these columns, based on their relevancy. The
final schema of the database is shown in {\bf listing \ref{schema}}
{\tt \small
\lstset{language = XML, caption={Database Schema}, label={schema}}
\begin{lstlisting}[float=*]
The Database has three tables:
1. mandb
2. mandb_meta
3. mandb_links

  COLUMN NAME       DESCRIPTION
  1. section        Section number
  2. name           The name of the man page  
  3. name_desc      Short description from NAME section  
  4. desc           Contains the content from DESCRIPTION section and any other
                    section which is not stored in a separate column.
  5. lib            The LIBRARY section
  6. return_vals    RETURN VALUES section
  7. env            ENVIRONMENT section
  8. files          FILES section
  9. exit_status    EXIT STATUS section
  10. diagnostics   DIAGNOSTICS section
  11. errors        ERRORS section
  12. md5_hash      An MD5 hash of the contents of the man page (UNIQUE).

(2) mandb_meta:
    This table maintains essential metadata of all the indexed man page files.

  COLUMN NAME       DESCRIPTION
  1. device         (dev_t)Logical device number from stat(2)
  2. inode          (ino_t)Inode number from stat(2)
  3. mtime          Last modification time from stat(2)
  4. file           Absolute path name (UNIQUE constraint)
  5. md5_hash       MD5 Hash of the man page file (UNIQUE)
  6. id             A unique integer ID referring to a page in mandb (PRIMARY KEY)

(3) mandb_links:
    This is an index of all the hard/soft links of the man pages. The list of
    the aliases is usually obtained from the multiple .Nm entries in the NAME
    section of a man page.

  COLUMN NAME       DESCRIPTION
  1. link           The name of the hard/soft link (UNIQUE index) 
  2. target         Name of the target page to which the link points  
  3. section        The section number
  4. machine        The machine architecture (if any) for which the page is
                    relevant.
\end{lstlisting}
}

\end{description}

@@

@@

\pagebreak
\section{Results}
{\bf Listing \ref{newUser}}, {\bf \ref{makeDirectory}}, and {\bf \ref{signal} } show results of some of the sample queries on new \textit{apropos}
to demonstrate how it performs as compared to the traditional version.
\endnote{Although the new \textit{apropos} by default returns 10
results for a query but in the listings the output has been
snipped to save space
} \\

{ \small
\lstset{language = XML, caption={Add new user}, label={newUser}}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%add new user%\normalfont\upshape''%
ssh-add(1)      adds private key identities to the authentication agent...on the command line.
If any file requires a passphrase, ssh-add asks for the passphrase from the user. The passphrase
is read from the user's tty. ssh-add retries the last passphrase if multiple identity files are
given...

@@

{\tt \small
\lstset{language = XML, caption={make directory}, label={makeDirectory}}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%make directory%\normalfont\upshape''%
make(1) maintain program dependencies
...CURDIR A path to the directory where make was executed. Refer to the description of PWD for
more details. MAKE The name that make was executed with argv[0]. For compatibility make also sets
.MAKE with the same value. The...

@@

{\tt \small
\lstset{language = XML, caption={signal number to string}, label={signal}}
\begin{lstlisting}[float=*]
$ apropos %\normalfont\upshape``%signal number to string%\normalfont\upshape''%
psignal(3)      system signal messages
...the signal number is not recognized sigaction(2) , the string Unknown signal is produced. The
psiginfo function produces the same output as the psignal function, only it uses the signal
number information from the si argument. The message strings can...

intro(2)        introduction to system calls and error numbers
...undefined signal to a signal(3) or kill(2) function). 23 ENFILE Too many open files in system.
Maximum number...shell. Pathname A path name is a NUL -terminated character string starting with
an optional slash, followed by zero or...

@@

\pagebreak
\section{Acknowledgement}
This project has been developed as part of \emph{Google Summer of Code 2011}
[17], thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. Thanks to \textbf{David Young} who was involved with
this project closely and offered useful guidance and encouragement throughout. A
special thanks goes to \textbf{Thomas Klausner} who helped in writing and
reviewing the man pages for this project. Thanks to \textbf{Petra Zeidler} for
administering the GSoC program for \textit{The NetBSD Foundation}. \\

@@

\end{thebibliography}
\theendnotes
